\textsf{Are large language models superhuman chemists?}

### 2024-03-29 src/tex/ms.tex
prepare addition of scaling analysis

% make the app somewhat more prominent

@@

    However, we still have only a very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of human chemists.

    We curated more than 7,000 question-answer pairs for a wide array of chemical sciences subfields, evaluated leading open and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average. 
    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident, misleading predictions, such as about chemicals' safety profiles.

@@

\paragraph{\enquote{Tiny} subset}
It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are more balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process).
We also used this subset to seed the web application for the human baseline study.

@@

Since many widely utilized systems only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
This is also important given the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems.
Such systems can augment the capabilities of \glspl{llm} through the use of external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
The text completions, however, are the system's final outputs, which would also be used in a real-world application. 
Hence, we use them for our evaluations.

\begin{figure}[!h]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered completely correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance. 
    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
    The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
    However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often not find the correct answer within the specified number of calls.
    In this case, we consider the answer as incorrect. 
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

@@

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}). We classified all questions in the \chembench corpus into these topics based on hand-crafted rules and classifier models operating on the question texts.
We then computed the percentage of questions the models or humans answered correctly for each topic.
\begin{figure}[!h]
    \centering
    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correctly by the models.
    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered completely correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}
In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). 
Thus, a larger colored area indicates a better performance. 
One can observe that this performance varies widely across models and topics. 
While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule). 
These findings also shine an interesting light on the value of textbook-inspired questions. 
A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.

@@

\section{Conclusions}
On the one hand, our findings underline the impressive capabilities of \glspl{llm} in the chemical sciences: Leading models outperform domain experts in specific chemistry questions on many topics. 
On the other hand, there are still striking limitations. 
For very relevant topics the answers models provide are wrong. 
On top of that, many models are not able to reliably estimate their own limitations.
Yet, the success of the models in our evaluations perhaps also reveals more about the limitations of the exams we use to evaluate models---and chemists---than about the models themselves.
For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning. 
Given that the models outperformed the average human in our study, we need to rethink how we teach and examine chemistry.
Critical reasoning is increasingly essential, and rote solving of problems or memorization of facts is a domain in which \glspl{llm} will continue to outperform humans.

>>>>

@@

    However, we still have only a very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve model performance and mitigate potential harms. 
    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs against the expertise of human chemists.

    We curated more than 7,000 question-answer pairs for a wide array of chemical sciences subfields, evaluated leading open and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average. 
    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident, misleading predictions, such as predictions related to chemicals' safety profiles.

@@

\paragraph{\enquote{Tiny} subset}
It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are more balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process). This \enquote{tiny} also provided the seed of the web application for our human baseline study.

@@

Since many widely utilized systems only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
This is also important given the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems.
Such systems can augment the capabilities of \glspl{llm} through the use of external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react, buehler2023generative}
In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
The text completions, however, are the system's final outputs, which would also be used in a real-world application. 
Hence, we use them for our evaluations.

\begin{figure}[!h]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered completely correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance. 
    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
    The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
    However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often fail to find the correct answer within the specified number of calls. If the \gls{llm} exceeds ten calls, we consider the answer as incorrect. 
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

@@

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}). We classified all questions in the \chembench corpus into these topics based on hand-crafted rules and classifier models operating on the question texts.
We then computed the percentage of questions the models or humans answered correctly for each topic.
\begin{figure}[!h]
    \centering
    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correctly by the models.
    The best score for every dimension is one (all questions answered correctly), and the worst is zero (no question answered completely correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that were answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}
In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). 
Thus, a larger colored area indicates a better overall performance across different substasks. 
One can observe that this performance varies widely across models and topics. 
While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule). 
These findings also shine an interesting light on the value of textbook-inspired questions. 
A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
On those questions, the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.

@@

\section{Conclusions}
On the one hand, our findings underline the impressive capabilities of \glspl{llm} in the chemical sciences: Leading models outperform domain experts in specific chemistry questions on many topics. 
On the other hand, there are still striking limitations. 
For very relevant topics the answers all models provide are wrong. 
On top of that, many models are not able to reliably estimate their own limitations.
Yet, the success of the models in our evaluations perhaps also reveals more about the limitations of the exams we use to evaluate models---and chemists---than about the models themselves.
For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning. 
Given that the models outperformed the average human in our study, we need to rethink how we teach and examine chemistry.
Critical reasoning is increasingly essential, and rote solving of problems or memorization of facts is a domain in which \glspl{llm} will continue to outperform humans.
